Type theories as quotient inductive-recursive types

### 2025-02-21 tex/main.tex
coquand universes

\subsection{The standard model} \label{subsec:std-model}
The standard model, justifying the soundness of our definition, is given by the same motives as those for its QIIT counterpart~\cite[Section~4]{Altenkirch2016a}:
\begin{align*}
  \Ctx^M\;\nameless          & \defeq \Set              & \Sub^M\;\Gamma^M\;\Delta^M\;\nameless & \defeq \Gamma^M \to \Delta^M \\
  \Ty^M\;\Gamma^M\;\nameless & \defeq \Gamma^M \to \Set & \Tm^M\;\Gamma^M\;A^M\;\nameless & \defeq (\gamma : \Gamma^M) \to A^M\;\gamma
\end{align*}
%with the additional methods to its QII counterpart:
%\begin{align*}
%  \blank\uparrow^M\blank & \defeq \lam{\sigma^M\,A^M\,\gamma} (\sigma^M\,(\pi_1\,\gamma) , \pi_2\,\gamma) & \\
%                         & : (\sigma^M : \Gamma^M \to \Delta^M) (A^M : \Delta^M \to \Set) \to \Sigma\;\Gamma^M\;(\lam{\gamma}A^M\,(\sigma^M\,\gamma)) \to \Sigma\,\Delta^M\,A^M \\
%  [\blank]_{\Tm}^M\blank & \defeq \lam{\sigma^M\,t^M\,\gamma} (t^M\,(\sigma^M\,\gamma))  \\
%                         & : (\sigma^M : \Gamma^M \to \Delta^M) (t^M : (\delta : \Delta^M) \to A^M\,\delta) \to (\gamma : \Gamma^M) \to A^M\,(\sigma^M\,\gamma)
%\end{align*}
%where $[\blank]_{\Tm}^M\blank$ is the same as the method for the explicit term substitution $[\blank]^M\blank$.
%
See our \Agda formalisation for details.
Because of the lack of transports in the syntax, many identities in the standard model hold strictly --- more so than in the QIIT standard model, where an extra step~\cite[Lemma~2.3.5]{UFP2013} is needed for reasoning about constant transports.

>>>>

\subsection{The standard model} \label{subsec:std-model}
The standard model of the type theory in \cref{subsec:SC+U+Pi}, justifying the soundness of our definition, is given by the same motives as those for its QIIT counterpart~\cite[Section~4]{Altenkirch2016a}:
\begin{align*}
  \Ctx^M\;\nameless          & \defeq \Set              & \Sub^M\;\Gamma^M\;\Delta^M\;\nameless & \defeq \Gamma^M \to \Delta^M \\
  \Ty^M\;\Gamma^M\;\nameless & \defeq \Gamma^M \to \Set & \Tm^M\;\Gamma^M\;A^M\;\nameless & \defeq (\gamma : \Gamma^M) \to A^M\;\gamma
\end{align*}
%with the additional methods to its QII counterpart:
%\begin{align*}
%  \blank\uparrow^M\blank & \defeq \lam{\sigma^M\,A^M\,\gamma} (\sigma^M\,(\pi_1\,\gamma) , \pi_2\,\gamma) & \\
%                         & : (\sigma^M : \Gamma^M \to \Delta^M) (A^M : \Delta^M \to \Set) \to \Sigma\;\Gamma^M\;(\lam{\gamma}A^M\,(\sigma^M\,\gamma)) \to \Sigma\,\Delta^M\,A^M \\
%  [\blank]_{\Tm}^M\blank & \defeq \lam{\sigma^M\,t^M\,\gamma} (t^M\,(\sigma^M\,\gamma))  \\
%                         & : (\sigma^M : \Gamma^M \to \Delta^M) (t^M : (\delta : \Delta^M) \to A^M\,\delta) \to (\gamma : \Gamma^M) \to A^M\,(\sigma^M\,\gamma)
%\end{align*}
%where $[\blank]_{\Tm}^M\blank$ is the same as the method for the explicit term substitution $[\blank]^M\blank$.
%
See our \Agda formalisation for details.
Because of the lack of transports in the syntax, many identities in the standard model hold strictly --- more so than in the QIIT standard model, where an extra step~\cite[Lemma~2.3.5]{UFP2013} is needed for reasoning about constant transports.
